---
title: "Project 3 Data Science Skills"
author: Umais Siddiqui,Neil Hwang, Michelle Mondy, Kalyanaraman Parthasarathy and
  Murali Kunissery
date: "October 17, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Rpubs link: http://rpubs.com/umais/Project3_Data607 | 

            http://rpubs.com/neilhwang/group2
            
            http://rpubs.com/mkunissery/321374 
            
            http://rpubs.com/mmondy/320996

GitHub link: https://github.com/umais/Data607-Project3

##Obejective

In this project our goal is to be able to answer the question 

**"Which are the most valued data science skills?"**

##Approach

In order to find the answer to our question we have researched a couple of job searching websites and decided to use CyberCoders and Indeed website to scrape the skillsets required for Jobs that had a title of Data Scientist.We will be using the rvest and MySQL libraries. rvest will be used to scrape and parse the HTML data.

```{r}
library(devtools)
library(RMySQL)
library(arules)
library(arulesViz)
library(ggplot2)
library(plyr)
```


##Database Schema

CREATE TABLE DataScienceJobs(


           JobId int auto_increment primary key,
           
           JobTitle nvarchar(255),
           
           JobLocation nvarchar(255),
           
           JobSalary nvarchar(255),
           
           Source nvarchar(255)
           
);


CREATE TABLE DataScienceSkills(

           SkillId int auto_increment primary key,
           
           JobId int,
           
           SkillName nvarchar(255)
           
);

##Data Collection

We seperated out the data collection part in in a R Script file called Project3.R

In that file we are scraping the data from CyberCoders and Indeed website for Jobs that have a title of data scientist and inserting it in to the tables created based on the above schema. Below is the link to that code.

https://github.com/umais/Data607-Project3/blob/master/Project3/Project3.R



##Retrieving Data from MySQL

We will be retrieving the data inserted in MYSQL from the scraping done from the R script file and performing some downstream analysis on the data.

```{r}
mydb = dbConnect(MySQL(), user='root', password='Welcome@1', dbname='project3', host='localhost')
#mydb = dbConnect(MySQL(), user='root', password='password', host= '127.0.0.1', port=3306)
#dbSendQuery(mydb, "CREATE DATABASE project3;")
#dbSendQuery(mydb, "USE project3")
  results = dbSendQuery(mydb, "SELECT j.JobTitle,j.JobLocation,s.SkillName FROM DataScienceJobs j INNER JOIN DataScienceSkills s ON j.JobId=s.JobId  ;")
  
  jobSkills=fetch(results, n=-1)
  
  head(jobSkills)
  
```
  
##More Analysis
  
  We can tell by the initial results from group by query that Python is the skill that is most valued.
  
```{r}
  rs = dbSendQuery(mydb, "SELECT SkillName,Count(1) as Total FROM DataScienceJobs j INNER JOIN DataScienceSkills s ON j.JobId=s.JobId GROUP BY SkillName Order By Total desc ;")
  
  df=fetch(rs, n=-1)
  

  
  head(df)
```

##Looking at Indeed Data

If we look at only Indeed data again we can see that Python , R and Machine Learning are among the top required skills.

```{r}
      rs = dbSendQuery(mydb, "SELECT SkillName,Source,Count(1) as Total FROM DataScienceJobs j INNER JOIN DataScienceSkills s ON j.JobId=s.JobId Where Source='Indeed' GROUP BY SkillName,Source Order By Total desc ;")
      
        IndeedDF=fetch(rs, n=-1)
        head(IndeedDF)
        
```

##Looking at CyberCoders Data 

Similarly in Cyber Coders data we see the same thing that Python,R,Machine Learning and Hadoop are among the top required skills for data scientist.

```{r}
    rs = dbSendQuery(mydb, "SELECT SkillName,Source,Count(1) as Total FROM DataScienceJobs j INNER JOIN DataScienceSkills s ON j.JobId=s.JobId Where Source='CyberCoders' GROUP BY SkillName,Source Order By Total desc ;")
      
        CyberCoderDF=fetch(rs, n=-1)
        head(CyberCoderDF)
  
  rs = dbSendQuery(mydb, "SELECT j.JobId,JobTitle,JobLocation,JobSalary,SkillName,Source FROM DataScienceJobs j INNER JOIN DataScienceSkills s ON j.JobId=s.JobId ;")
  
  AllJobs=fetch(rs, n=-1)
  
head(AllJobs)
```

##Association Analysis

To explore the data further, we perform an association analysis, which is one of the more popular unsupervised machine learning algorithms, using the package *arules*.
To begin, we preoprocess the data to list the skills by specific jobs.

```{r}
df <- data.frame(matrix(ncol = 2, nrow = nrow(AllJobs)))
df[,1] <- factor(AllJobs[,"JobId"])
df[,2] <- factor(AllJobs[,"SkillName"])
temp <- df[,c(1,2)]
first_item <- ddply(temp, .(X1), function(x) x[1, ])
temp2 <- merge(x = temp, y = first_item, by = "X1", all.x = TRUE)
data <- temp2[duplicated(temp2$X1),]
data$X1 <- data$X2.y
data$X2.y <- NULL
names(data) <- c("X", "Y")
m <- as.matrix(data)
l <- lapply(1:nrow(m), FUN = function(i) (m[i, ]))
```

Now, we convert the list into transactions that *arules* can work with in forming aprior rules to identify the most common associations that tend to occur together among skills. 
```{r}
transactions <- as(l, "transactions")
itemsets <- apriori(transactions, parameter = list(target = "frequent",
    supp=0.001, minlen = 2, maxlen=6))
```

This frequency plot confirms our findings earlier that among the most demanded skill of data scientists are python, machine learning, R, data mining, and hadoop.
```{r}
#Top 30 most frequently occurring skills
itemFrequencyPlot(transactions, topN=30)
quality(itemsets)$lift <- interestMeasure(itemsets, measure="lift", trans = transactions)

#Top 30 associations
inspect(head(sort(itemsets, by = "lift"), n=30))

#Visualization of top associations and skills
plot(head(sort(itemsets, by = "lift"), n=30), method = "graph", control=list(cex=.8))
```


##Conclusion

Based on the data collected from CyberCoders and Indeed we can see after doing some analysis that the most valuable skills for a data scientist are Python, R , Machine learning and Hadoop.

