---
title: "Project 3 Group 2"
author: "Michelle Mondy"
date: "October 19, 2017"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Overview
For this project, we will use data to answer the question "What are the most valued data science skills?" For our approach, we have decided to scrape data from both Cybercoders and Indeed job listings to see which skills are most sought after when looking to hire a Data Scientist.

##Load Required Packages
```{r, comment=NA, results='hide', warning=FALSE} 
library(rvest)
library(stringr)
library(RMySQL)
library(tidyr)
library(dplyr)
```


##Grab data from Indeed
In order to scrape data from Indeed, we identified tags using the SelectorGadget. We decided to use the the first 10 pages of job listings results assuming they would be the most relevant.

```{r}
getData <- function(url){
IndeedJobTitle<-url %>%
  read_html() %>%
  html_nodes("#resultsCol .jobtitle") %>%
  html_text()
IndeedJobTitle_clean <<- str_extract(IndeedJobTitle, "[[:alpha:]., /-]{2,}")

IndeedJobLocation<-url %>%
  read_html() %>%
  html_nodes(".location") %>%
  html_text()
IndeedJobLocation_clean <<- str_extract(IndeedJobLocation, "[[:alpha:], ]{2,}")

IndeedSkills<-url %>%
  read_html() %>%
  html_nodes(".experience") %>%
  html_text()
IndeedSkills_clean <<- str_extract(str_sub(IndeedSkills, start=22), "[[:alpha:], ]{2,}")
}

getData("https://www.indeed.com/jobs?q=data+scientist&l=")
page1 <- data.frame(IndeedJobTitle_clean,IndeedJobLocation_clean,IndeedSkills_clean)
getData("https://www.indeed.com/jobs?q=Data+Scientist&start=10")
page2 <- data.frame(IndeedJobTitle_clean,IndeedJobLocation_clean,IndeedSkills_clean)
getData("https://www.indeed.com/jobs?q=Data+Scientist&start=20")
page3 <- data.frame(IndeedJobTitle_clean,IndeedJobLocation_clean,IndeedSkills_clean)
getData("https://www.indeed.com/jobs?q=Data+Scientist&start=30")
page4 <- data.frame(IndeedJobTitle_clean,IndeedJobLocation_clean,IndeedSkills_clean)
getData("https://www.indeed.com/jobs?q=Data+Scientist&start=40")
page5 <- data.frame(IndeedJobTitle_clean,IndeedJobLocation_clean,IndeedSkills_clean)
getData("https://www.indeed.com/jobs?q=Data+Scientist&start=50")
#page6 <- data.frame(IndeedJobTitle_clean,IndeedJobLocation_clean,IndeedSkills_clean)
getData("https://www.indeed.com/jobs?q=Data+Scientist&start=60")
page7 <- data.frame(IndeedJobTitle_clean,IndeedJobLocation_clean,IndeedSkills_clean)
getData("https://www.indeed.com/jobs?q=Data+Scientist&start=70")
page8 <- data.frame(IndeedJobTitle_clean,IndeedJobLocation_clean,IndeedSkills_clean)
getData("https://www.indeed.com/jobs?q=Data+Scientist&start=80")
page9 <- data.frame(IndeedJobTitle_clean,IndeedJobLocation_clean,IndeedSkills_clean)
getData("https://www.indeed.com/jobs?q=Data+Scientist&start=90")
page10 <- data.frame(IndeedJobTitle_clean,IndeedJobLocation_clean,IndeedSkills_clean)


indeed_data <- rbind(page1, page2, page3, page4, page5, page7, page8, page9, page10)
```

##Creating the MySQL tables
To follow the same structure as Umais, I have created the following 2 tables in MySQL. Note: The password will need to be updated in the connection statement. 

```{r}
mydb = dbConnect(MySQL(), user='root', password='sqlpass', dbname='project3', host='localhost')

dbGetQuery(mydb, "DROP TABLE IF EXISTS IndeedDataScienceJobs;" )
dbGetQuery(mydb, "DROP TABLE IF exists IndeedDataScienceSkills;" )
dbGetQuery(mydb, "CREATE TABLE IndeedDataScienceJobs(
  JobId int auto_increment primary key,
  JobTitle nvarchar(255),
  JobLocation nvarchar(255)
  );" )

dbGetQuery(mydb, "CREATE TABLE IndeedDataScienceSkills(
  SkillId int auto_increment primary key,
  JobId int,
  SkillName nvarchar(255)
  );" )
```

##Transforming the Data to make it more usable

Since we have the data in a data frame, but it is not very usuable yet, we would like to insert it into a separate tables in MYSQL and use sql to transform it. Once the data is in the format that fits the tables created above, it will be insert into those tables. 

```{r}
dbWriteTable(mydb, "indeed_data", indeed_data, overwrite = TRUE)

#This query creates a separate field (skill1-12) for each of the skills listed in the "IndeedSkills_clean" column. It also strips out the white space.

trans_query <- 
"SELECT 
IndeedJobTitle_clean
,IndeedJobLocation_clean
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',12 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',11 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',12 ), ',', -1)) else null 
end as skill12
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',11 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',10 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',11 ), ',', -1)) else null 
end as skill11
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',10 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',9 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',10 ), ',', -1)) else null 
end as skill10
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',9 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',8 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',9 ), ',', -1)) else null 
end as skill9
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',8 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',7 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',8 ), ',', -1)) else null 
end as skill8
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',7 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',6 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',7 ), ',', -1)) else null 
end as skill7
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',6 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',5 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',6 ), ',', -1)) else null 
end as skill6
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',5 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',4 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',5 ), ',', -1)) else null 
end as skill5
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',4 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',3 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',4 ), ',', -1)) else null 
end as skill4
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',3 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',2 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',3 ), ',', -1)) else null 
end as skill3
,case when 
substring_index ( substring_index ( IndeedSkills_clean,',',2 ), ',', -1) <> substring_index ( substring_index ( IndeedSkills_clean,',',1 ), ',', -1) 
then trim(substring_index ( substring_index ( IndeedSkills_clean,',',2 ), ',', -1)) else null 
end as skill2
,trim(substring_index ( substring_index ( IndeedSkills_clean,',',1 ), ',', -1)) as skill1
FROM project3.indeed_data;"

wide_data <- dbGetQuery(mydb, trans_query)
head(wide_data)


#Write the data to MYSQL and then use that table to Insert into the IndeedDataScienceJobs table
dbGetQuery(mydb, "DROP TABLE IF EXISTS indeed_wide_data;" )
dbWriteTable(mydb, "indeed_wide_data", wide_data, overwrite = TRUE)
dbGetQuery(mydb, "INSERT INTO IndeedDataScienceJobs(JobId, JobTitle,JobLocation) 
                  Select row_names,IndeedJobTitle_clean, IndeedJobLocation_clean
                  from project3.indeed_wide_data")

wide_data$ID <- seq.int(nrow(wide_data))
#Using the Tidyr package, we will rehape the data
skills <- gather(wide_data,"n", "skill", 3:14)
#Using dplyr we will subset the data to exclude null rows
skills2 <- filter(skills, !is.na(skills$skill)) %>% 
  select(ID, skill) %>% 
  arrange(ID)
skills2$skill_ID <- seq.int(nrow(skills2))

#Write the data to MYSQL and then use that table to Insert into the IndeedDataScienceSkills table
dbGetQuery(mydb, "DROP TABLE IF EXISTS skills_data;" )
dbWriteTable(mydb, "skills_data", skills2, overwrite = TRUE)
dbGetQuery(mydb, "INSERT INTO IndeedDataScienceSkills(SkillId, JobId, SkillName) 
                  Select skill_ID,ID, skill
                  from project3.skills_data")
```


##Analysis
Now that we have the data in a usable format, we can do some analysis. Let look at the Top 15 most valued Data Science skills.

```{r}
Most_valued_skills <- count(skills2, skill) %>% arrange(desc(n))
Most_valued_skills


top_15_skills <- dbGetQuery(mydb, "select SkillName, count(*) as n from project3.indeeddatascienceskills
                                  group by SkillName
                                  order by 2 desc
                                  Limit 15;")



library(ggplot2)
ggplot(data=top_15_skills, aes(x=SkillName,y=n)) +
  geom_bar(position="dodge",stat="identity") + 
  coord_flip() +
  ggtitle("Most Valued Data Science Skills") +
  scale_x_discrete(
    limits= c("Scala", "Sas", "Data Warehouse", "MATLAB","AI","Natural Language Processing", "Java","Data Mining","Spark", 
              "Hadoop", "Data Science", "Big Data", "Python","R", "Machine Learning")
  ) +
theme(
  axis.title.y=element_blank(),
  axis.title.x=element_blank()
  )
```




